---
title: "RiTMIC: RegulatIon of Tumor MIcroenvironment Composition"
subtitle: "Part 2 : Deconvolution with medepir package"
author: "Magali Richard, Clementine Decamps, Fabien Quinquis"
contact: 
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{deconvolution}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=TRUE, results="verbatim", dpi=75)
layout(1, respect=TRUE)
```


# Introduction

`RiTMIC` (**R**egulat**I**on of **T**umor **MI**croenvironment **C**omposition) is an open-access R package to study the link between the gene expression in a tumor and the microenvironment composition. 

**To do** : describe the pipeline's steps : 1. simulation of complex D matrix / 2. deconvolution with medepir / 3. identification and extraction of the cancer type / 4. Penda / 5. res analysis.

`RiTMIC` Part 2: **Deconvolution of D matrices**

#Load data

**To do** : describe data obtained in simulation

```{r, label="load_data"}
D_met = readRDS("tmp_ritmic/D_met.rds")
D_rna = readRDS("tmp_ritmic/D_rna.rds")
```

#Step 1 : deconv methylation 

**To do** : describe differents steps : 
Feature selection, deconv avec EDEc (explain edec)

```{r, label="Deconv meth"}
D_FS = medepir::feature_selection(D = D_met, number = 2000)
results_edec = medepir::Edec(D_FS, nbcell = 4, infloci = rownames(D_FS))

if(!file.exists("tmp_ritmic/res_edec.rds")){
  saveRDS(results_edec, file = "tmp_ritmic/res_edec.rds")
}
```

#Step 1 : deconv RNAseq 

**To do** : describe differents steps : 
EDec step 2 , take D rnases and A computed on D meth to obtain T meth 

```{r, label="Deconv rnaseq"}
prop_cell_type = t(results_edec$A)
res_edec2 = EDec::run_edec_stage_2(D_rna, prop_cell_type)
T_res = res_edec2$means
colnames(T_res) = 1:4

if(!file.exists("tmp_ritmic/res_edec2.rds")){
  saveRDS(T_res, file = "tmp_ritmic/res_edec2.rds")
}
```
