---
title: "RiTMIC: RegulatIon of Tumor MIcroenvironment Composition"
subtitle: "Step 4: Analysis of gene deregulation with `penda`"
author: "Magali Richard, Clementine Decamps, Fabien Quinquis"
contact: 
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{penda_analysis}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=TRUE, results="verbatim", dpi=75)
layout(1, respect=TRUE)
```

# Introduction 

`RiTMIC` (**R**egulat**I**on of **T**umor **MI**croenvironment **C**omposition) is an open-access R package to study the link between the gene expression in a tumor and the microenvironment composition. 

**To do** : describe the pipeline's steps : 1. simulation of complex D matrix / 2. deconvolution with medepir / 3. identification and extraction of the cancer type / 4. Penda / 5. res analysis.

`RiTMIC` **"Step 4: Analysis of gene deregulation  with `penda` **

`penda` (**PE**rso**N**alized **D**ifferential **A**nalysis ) is an open-access R package that detects gene deregulation in individual samples compared to a set of reference, control samples. 

# Dataset

**To do** : describe data 

```{r, label="load_data"}
D_cancer = readRDS("tmp_ritmic/D_cancer.rds")
```

# Penda Method

`penda` performs a 3-steps analysis: 
  1. Data filtering
  2. Relative gene ordering
  3. Differential expression testing

**To do** : describe that we used as controls and what user has to used 
Described make dataset

## Data filtering
```{r echo=FALSE, label="data filtering"}
penda_data_case <- D_cancer
penda_data_ctrl <- penda::penda_data_ctrl[1:1000, 1:10]

Penda_dataset = penda::make_dataset(penda_data_ctrl, penda_data_case, detectlowvalue = F)
data_ctrl = Penda_dataset$data_ctrl
data_case = Penda_dataset$data_case
```


## Relative gene ordering

**To do** : explain what we make and what user has to make (simu penda)

```{r, label= "penda::compute_lower_and_higher_lists"}
s_max = 30
L_H_list = penda::compute_lower_and_higher_lists(data_ctrl, threshold = 0.99, s_max = s_max)
L = L_H_list$L
H = L_H_list$H
```

## Differential expression testing

**To do** : explain what we make and what user has to make (simu penda)

```{r, label="penda::penda_test", results="hide"}
threshold = 0.4
iterations =  20

penda_res = penda::penda_test(samples = data_case, 
                              controls = data_ctrl,
                              iterations =  iterations, 
                              L_H_list =  L_H_list, 
                              threshold = threshold)

if(!file.exists("tmp_ritmic/penda_res.rds")){
  saveRDS(penda_res, file = "tmp_ritmic/penda_res.rds")
}

```
