---
title: "RiTMIC: RegulatIon of Tumor MIcroenvironment Composition"
subtitle: "Part 5 : Analysis of outputs"
author: "Magali Richard, Clementine Decamps, Fabien Quinquis"
contact: 
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=TRUE, results="verbatim", dpi=75)
layout(1, respect=TRUE)
```


# Introduction

`RiTMIC` (**R**egulat**I**on of **T**umor **MI**croenvironment **C**omposition) is an open-access R package to study the link between the gene expression in a tumor and the microenvironment composition. 

**To do** : describe the pipeline's steps : 1. simulation of complex D matrix / 2. deconvolution with medepir / 3. identification and extraction of the cancer type / 4. Penda / 5. res analysis.

`RiTMIC` Part 5: **Res analysis**


# Informations

If curves of distance methods has a higher True Positive Rate (TPR) combined a lower False Positive Rate(FPR) than correlation curve, __PenDA__ improves the model.  

Three data input are required to perform the statistical analysis :

- penda output : the down and up matrices of dimension: n*g 
  - g: gene names 
  - n: the deregulation status of the cancer compared to the control
    - -1: Down-regulated gene 
    - 0: Same expression
    - +1: Up-regulated gene 

- Matrix A of cell lines proportion distribution simulated in `step_1_simulation`
  - proportion: k*n
  - n: patient 
  - k: proportion of the different cell lines
  
- Matrix T of reference cell lines simulated for the example in `step_1_simulation`
  - proportion: ge x nc
  - ge: gene expression
  - nc: number of cell lines 
  
  
#Method 

## Step 1: Loading required dataset and apply RiTMIC::pre_treat function

**To do** : describe data and pre treat function

```{r, label="load_data"}
D_cancer = readRDS("tmp_ritmic/D_cancer.rds")
A_calc = readRDS("tmp_ritmic/res_edec.rds")$A
res_penda = readRDS("tmp_ritmic/penda_res.rds")
genes_dereg = readRDS("tmp_ritmic/T_tum_dereg.rds")$g_dereg

res_penda_f = Ritmic::pre_treat(penda_res = res_penda, thres_p = 5)
```

## Step 2: Compute metrics on penda and on gene expression

**To do** : describe this step

```{r, label="compute metrics "}
res_dereg <- Ritmic::calc_dist(binary_penda = res_penda_f, A = A_calc)
res_corr <- Ritmic::calc_corr(D_cancer = D_cancer, A = A_calc)
```

## Step 3: ROC curve

**To do** : describe this step

Find where is the ME cell type in the calc matrices

```{r, label="ME identification"}
A_ref = readRDS("tmp_ritmic/A_matrix.rds")
rownames(A_ref) = c("", "", "immune","")
cor(t(A_ref), t(A_calc))
pos_ME = which.max(cor(t(A_ref), t(A_calc))[3,])
```

**To do** : describe this step

```{r,label="RiTMIC::plot_res"}
#To focus on the cell type 3 
corr_ME = res_corr[res_corr$type == pos_ME, ]
dist_ME = res_dereg[res_dereg$type == pos_ME, ]

pvalues = c(0, 0.00005, 0.0001, 0.0005, 0.001, 0.0025, 0.005, 0.0075, 0.01, 0.02, 0.03, 0.04, 0.05, 0.1, 0.15, 0.2, 0.25,  0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1)

Ritmic::plot_res(corr_res = corr_ME, dist_res = dist_ME, genes_dereg = genes_dereg, pvalues, graph_title = "ROC curves to test the model improvement induce by PenDA")
```

# Session Information

```{r, results="verbatim"}
sessionInfo()
```