---
title: "RiTMIC: RegulatIon of Tumor MIcroenvironment Composition"
subtitle: "Step 1 : Increase your RNAseq datasets with a new simulated adapted dataset derivated from yours"
author: "Magali Richard, Clementine Decamps, Florent Chuffart, Fabien Quinquis, Daniel Jost"
contact: 
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=TRUE, results="verbatim", dpi=75)
layout(1, respect=TRUE)
```

# Introduction

`RiTMIC` (**R**egulat**I**on of **T**umor **MI**croenvironment **C**omposition) is an open-access R package that simulates the necessary tools to perform `PenDA` analysis, perform analysis with `PenDA` and apply visualization tools to the `PenDA` output. 

`RiTMIC` Step 1: **Simulation of inputs required to the R package analysis: `penda`**

# Generation of input files for penda analysis

`simulation` has been developed to simulate an increased dataset that fits with your RNAseq gene distribution dataset and the proportion of cell lines distribution per sample. This functionality requires the different cell proportions distributions.

The output matrix of gene expression per sample is required to perform. This matrix can be directly used in `PenDA`.

`PenDA` analysis is an open-access R package that detects gene deregulation in individual samples by relative gene ordering and differential expression testing. This analysis will be performed on step 2 of the pipeline.

## Explanation of dataset

`data_ctrl` is a data matrix containing the normalized counts of the control sample. 
The names of the matrix correspond to the gene_symbol.

`data_case` is a data matrix containing the normalized counts of each tumor sample. 
The rownames of the matrix correspond to the gene_symbol, the colnames indicate the sample ID.

## Increase your dataset with a randomized gene expression distribution

```{r, label="RiTMIC::simu_A"}
cell_type_proportions = c(1.5, 4.5, 1, 3)
tumorous_sample_number = 40
matrix_A = RiTMIC::simu_A(tumorous_sample_number,cell_type_proportions)
```

The function `simu_A` simulates the matrix of proportion of cell lines distribution per tumor. Be careful, the `tumorous_sample_number` needs to be the new wanted tumor number for `simu_T_cancer` function

## Generation of the matrix T : gene expressions per cell line  

{r, label="RiTMIC::simu_T_cancer"}
tumor_dataset <- "Your RNAseq dataset"
tumorous_sample_number = 40
matrix_T = RiTMIC::simu_T_cancer(tumor_dataset,tumorous_sample_number)


New simulated RNAseq samples can be obtained from your RNAseq dataset with `simu_T_cancer`: for each gene, novel gene expressions will be assigned by a random choice inside the normal distribution of the actual gene expression   


